An efficient approach to approximating the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-08-29 papers/pair-correlation/paper.tex
Changed the end of section IV in pair paper. Tested on knightley. Took 1:24.97.

\newcommand{\fixme}[1]{\red{[#1]}}

@@

The CVA requires the radial distribution function expressed as a function of $r$
and $g_\sigma$.  One approach to constructing $g(r;g_\sigma)$ is to
compute $n$ from $g_\sigma$ (using e.g. the Carnahan-Starling formula
for $g_\sigma$)
and then find $g(r;n)$ using either
some approximate functional form (e.g. Percus-Yevick) or Monte Carlo
simulation results.  In this paper we use a fit to Monte Carlo results
for the radial distribution function $g(r)$ which facilitates
efficient computation.  When evaluating the approximations of earlier
works described in Section~\ref{sec:gn}, we use interpolated Monte
Carlo results for $g(r;n)$.

@@

\section{Making the CVA efficient}
The existing approaches to approximating the pair distribution
function outlined in Section~\ref{sec:gn} have not been widely used in
the construction of density functionals based on thermodynamic
perturbation theory, largely due to their computational complexity.
While our CVA provides only an incremental improvement in accuracy,
its construction enables significant gains in computational
efficiency, allowing for practical application in density functionals.
We achieve this gain by developing a \emph{separable} fit to the radial
distribution function of the hard-sphere fluid (see
Section~\ref{sec:separable-fit} for details).  This separable fit is
of the form
\begin{align}
  g(r_{12}; g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
\end{align}
where the notable aspect is that the radial distribution function is
written as a sum of terms that are each a simple product of a function
of radius with a function of $g_\sigma$.  This enables us to write
integrals---such as Eq.~\ref{eq:mean-field}---that are linear in the
pair distribution function as a summation of fixed-kernel
convolutions, which may be efficiently computed using Fast Fourier
Transforms (FFTs) in $\mathcal{O}(N\log N)$ time.

>>>>

\newcommand{\fixme}[1]{\red{[#1]}}
\newcommand{\newstuff}[1]{\blue{#1}}

@@

The CVA requires the radial distribution function expressed as a
function of $r$ and $g_\sigma$.  \fixme{One approach to constructing
  $g(r;g_\sigma)$ is to compute $n$ from $g_\sigma$ (using e.g. the
  Carnahan-Starling formula for $g_\sigma$) and then find $g(r;n)$
  using either some approximate functional form (e.g. Percus-Yevick)
  or Monte Carlo simulation results.}  \newstuff{In this paper we
  construct a functional form for $g(r, g_\sigma)$ that allows for
  greater computational efficiency. In section \ref{sec:efficient} we
  discuss the general form that allows for this efficiency and in
  section \ref{sec:separable-fit} we detail the fit to Monte Carlo
  results.}  When evaluating the approximations of earlier works
described in Section~\ref{sec:gn} \newstuff{for comparison}, we use interpolated Monte Carlo
results for $g(r;n)$.

@@

\section{Making the CVA efficient}
\label{sec:efficient}
The existing approaches to approximating the pair distribution
function outlined in Section~\ref{sec:gn} have not been widely used in
the construction of density functionals based on thermodynamic
perturbation theory, largely due to their computational complexity.
While our CVA provides only an incremental improvement in accuracy,
its construction enables significant gains in computational
efficiency, allowing for practical application in density functionals.
We achieve this gain by developing a \emph{separable} fit to the radial
distribution function of the hard-sphere fluid (see
Section~\ref{sec:separable-fit} for details).  This separable fit is
of the form
\begin{align}
  g(r; g_\sigma) &= \sum_{i} a_i(r) b_i(g_\sigma)
  \\\text{\newstuff{removed subscript $_{12}$.}}
\end{align}
where the notable aspect is that the radial distribution function is
written as a sum of terms that are each a simple product of a function
of radius with a function of $g_\sigma$.  This enables us to write
integrals---such as Eq.~\ref{eq:mean-field}---that are linear in the
pair distribution function as a summation of fixed-kernel
convolutions, which may be efficiently computed using Fast Fourier
Transforms (FFTs) in $\mathcal{O}(N\log N)$ time.
